An Introduction to Return Oriented Programming

### 2023-02-17 paper.tex
changes

%%%% 6. ABSTRACT %%%%
\begin{abstract}
  Return Oriented Programming is a buffer overflow exploitation technique developed in 2007. Under certain circumstances it can provide arbitrary code execution on assembly level. Luckily the development of binary exploitation protections like stack canaries, bounds checking and ASLR made the effort required for developing an attack for that specific target and attacking the target too high to make it practicable in most cases, however, old soft-/hardware are still vulnerable and the mentioned protections have shown to be vulnerable themselves, with high enough reward ROP may be a devastating technique for black-hats.  
\end{abstract}

%%%% 7. PAPER CONTENT %%%%
\section{Introduction}
Return Oriented Programming, abbreviated ROP is a type of buffer overflow attack that has been published in 2007 by Hovav Shacham.~\cite{ropfirstpaper} and has become a widely known buffer overflow technique since. It has been developed to circumvent the NX-BIT protection that protects the stack from being executed. At the time of writing this paper modern techniques like stack carnaries and ASLR make these attacks hard and very time consuming on modern systems. That is not to say ASLR and stack canaries can not be broken by bruteforcing or side channels. Since there are millions of running systems with old hard-, firm- and software that is possibly vulnerable to these kinds of attacks it is still relevant to this day. The main idea in ROP is based on chaining return addresses to code just before a return and therefore allowing almost arbitrary cpu instructions to be chained.

@@

This command produces an output with results similar to this.

@@

\section{Theory}
\subsection{Stack}
The following graphic~\cref{fig:stack} is an illustration of how the stack changes when injecting the payload. The buffer first has to be filled. In binary exploitation the letter \Verb+A+ is used for that most of the time, it has an easy to identify hexadecimal value of \bltInlineVerb{0x41}. It is important to note that without any special compiler options the stack will be aligned in \bltInlineVerb{dword}'s/16 Byte blocks. because of that the buffer has to be filled with more bytes than the buffer holds if $s \mod 16 \neq 0$ holds true, s being the buffer size in Bytes.
\newpage
\begin{figure}[h!]
  \includegraphics[width=0.79\textwidth]{stackropoffsec.png}
  \caption{The stack when injecting the payload}
  \label{fig:stack}
\subsection{ROP Runtime Behaviour}
The following graphic~\cref{fig:executionatruntime} illustrates how the gadgets get executed once the instruction pointer \bltInlineVerb{eip} points to the \bltInlineVerb{ret} in \bltInlineVerb{main}. Once this happens the execution gets redirected to the first gadget and executes the instructions in it. As soon as \bltInlineVerb{eip} points to the \bltInlineVerb{ret} in the 1st gadget the address of the 2nd gadget is \bltInlineVerb{pop}'d into \bltInlineVerb{eip} and execution continues there, from there the same thing happens again until execution reaches the end of the last gadget.
\begin{figure}[h!]
  \includegraphics[width=0.95\textwidth]{Ropchaineffect.png}
  \caption{The stack when injecting the payload}
  \label{fig:executionatruntime}
\section{Attack}

@@

\paragraph{Phases}

@@

\subsection{The attack}

@@

The segments in a binary can be read only or writable. It is possible to determine wether a segment is read only with \bltInlineVerb{objdump -h}. However, the following~\cref{command:finddatasegment} bash command can be used to find the address of the data segment. The data segment contains static and global variables. Since the target program does not have any global or static variables we can override this segment with arbitrary character sequences. In 

>>>>

%%%% 6. ABSTRACT %%%%
\begin{abstract}
  In this paper we introduce the concept, demonstrate an attack and talk about the different protection mechanisms of Return Oriented Programming. ROP is a buffer overflow exploitation technique developed in 2007. Under certain circumstances it can provide arbitrary code execution on assembly level. Modern 64 Bit binaries are decently secure against ROP with ASLR and stack protections enabled. 32 Bit binaries or binaries compiled for non PC systems may not provide the same protection though and may be vulnerable to ROP.
\end{abstract}

\section{Introduction}

%%%% 7. PAPER CONTENT %%%%
\section{Introduction}
Return Oriented Programming, abbreviated ROP is a type of buffer overflow attack that has been published in 2007 by Hovav Shacham.~\cite{ropfirstpaper} and has become a widely known buffer overflow technique since. It has been developed to circumvent the NX-BIT protection that protects the stack from being executed. The general consensus is that modern binaries are practically not vulnerable to buffer overflow attacks, but there is a lot of research surrounding breaking of these security measures that shows practical strength of these security measures does not equal the theoretical strength due to side channels, bugs or other exploits.~\cite{aslr} With high enough reward ROP may be a devastating technique for black-hats, because of that it is important to raise awareness about binary exploitation generally and ROP. Because of that this paper will demonstrate the underlying theory and demonstrate it with an attack on a vulnerable binary.

@@

This command produces an output with results similar to this~\cref{outputropgadget}.

@@

\section{Theory}
\subsection{Stack}
The following graphic~\cref{fig:stack} is an illustration of how the stack changes when injecting the payload. The buffer first has to be filled. In binary exploitation the letter \Verb+A+ is used for that most of the time, it has an easy to identify hexadecimal value of \bltInlineVerb{0x41}. It is important to note that without any special compiler options the stack will be aligned in \bltInlineVerb{dword}'s/16 Byte blocks. because of that the buffer has to be filled with more bytes than the buffer holds if $s \mod 16 \neq 0$ holds true, s being the buffer size in Bytes.
\begin{figure}[h!]
  \centering
  \includegraphics[width=0.79\textwidth]{stackropoffsec.png}
  \caption{The stack when injecting the payload}
  \label{fig:stack}
\end{figure}
\\
\subsection{ROP Runtime Behaviour}
The following graphic~\cref{fig:executionatruntime} illustrates how the gadgets get executed once the instruction pointer \bltInlineVerb{eip} points to the \bltInlineVerb{ret} in \bltInlineVerb{main}. Once this happens the execution gets redirected to the first gadget and executes the instructions in it. As soon as \bltInlineVerb{eip} points to the \bltInlineVerb{ret} in the 1st gadget the address of the 2nd gadget is \bltInlineVerb{pop}'d into \bltInlineVerb{eip} and execution continues there, from there the same thing happens again until execution reaches the end of the last gadget.
\begin{figure}[h!]
  \centering
  \includegraphics[width=0.95\textwidth]{Ropchaineffect.png}
  \caption{The stack when injecting the payload}
  \label{fig:executionatruntime}
\end{figure}

\section{Attack: Opening a Shell}

@@

@@

\subsection{Opening a Shell}

@@

The segments in a binary can be read only or writable. It is possible to determine wether a segment is read only with \bltInlineVerb{objdump -h}. However, the following~\cref{command:finddatasegment} bash command can be used to find the address of the data segment. The data segment contains static and global variables. Since the target program does not have any global or static variables we can override this segment with arbitrary character sequences.
